---
title: "Differential Analysis Results"
author: "Paulina Urbanaviciute"
date: "2024-04-03"
output:
  rmarkdown::html_document:
    theme: flatly
    highlight: tango
    code_folding: show
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
    number_sections: no
    self_contained: yes
    df_print: "paged"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(purrr)
library(stringr)
library(ComplexHeatmap)
library(clusterProfiler)
library(enrichplot)
library(ggplot2)
library(dplyr)
library(ggrepel)
library(ggvenn)
library(RColorBrewer)
library(knitr)

# set working directory
workdir="/Volumes/pu1918/ephemeral/AS_astrocytes"
# set seed
set.seed(123)
```


# Gene lists {.tabset} 
View results table from gene differential analysis.

differential_analysis_BMP4_CNTF
```{r}
# load de outputs
res.file<-paste0(workdir, "/analysed_data/differential_analysis_BMP4_CNTF.xlsx")
dge_res_sheets <- readxl::excel_sheets(res.file)[-c(1:4)]
dge_res_sheets <- data.frame(sheets = dge_res_sheets, row.names = dge_res_sheets)
dge_res_sheets_key <- readxl::read_xlsx(res.file, sheet = "Comparison Key")

# replace long sheet name from the keys
dge_res_sheets[dge_res_sheets_key$Key, ] <- dge_res_sheets_key$Comparison

# select comparisons of interest 
dgeBulk <- imap(rownames(dge_res_sheets), ~readxl::read_xlsx(res.file, sheet = .x)) %>%
    setNames(nm = dge_res_sheets$sheets)
# dgeBulk_sel <- dgeBulk[c("BMP4_diff_3w - progenitor|untreated",
#                          "CNTF_diff_3w - progenitor|untreated")]
dgeBulk_sel <- dgeBulk[c(dge_res_sheets_key$Comparison)]
# select relevant comparisons
dgeBulk_sel = dgeBulk_sel[-c(6,8,14)]
```

```{r, results='asis'}
# for each comparison, return a table of top hits
deTopList <- c()
for (i in 1:length(dgeBulk_sel)) {
    # select item from list
    de = dgeBulk_sel[[i]]
    name=names(dgeBulk_sel)[i]
    deTop = de %>% arrange(pvalue) %>% slice_head(n=100)
    deTopList <- c(deTopList, list(deTop ))
  }
```


## BMP4_diff_3w - progenitor|untreated
```{r}
i=2 # "BMP4_diff_3w - progenitor|untreated"
df = deTopList[[i]]

print(names(deTopList[i]))
data.frame(df)
```
## CNTF_diff_3w - progenitor|untreated
```{r}
i=6 # "CNTF_diff_3w - progenitor|untreated"
df = deTopList[[i]]

print(names(deTopList[i]))
data.frame(df)
```

# Gene counts {.tabset} 
```{r}
countSig <- function(i) {
    # select item from list
    de = dgeBulk_sel[[i]]
    name=names(dgeBulk_sel)[i]

    # count significant up and down DEGs
    deCount = de %>% filter(sig=="<= 0.05") %>% group_by(class) %>% summarise(count=n())

    return(deCount)
}

# apply function to list of comparisons
clist = 1:length(dgeBulk_sel)
outs = lapply(clist, countSig)
names(outs) <- names(dgeBulk_sel)
```

View counts of significant differential analysis hits.

```{r, message=FALSE, warning=TRUE, fig.height=10, fig.width=10, echo=FALSE, results='asis',fig.keep='all'}

plotl=outs

for (i in 1:length(dgeBulk_sel)) {

    name = names(dgeBulk_sel)[[i]]
    cat('\n')
    cat('## ', name, '   \n')
    cat('\n')

    data.frame(plotl[[i]])

    cat('\n')

}
```

differential_analysis_wo2W_BMP4_CNTF_wo2W 
```{r}
# # load differentiated BMP4 and CNTF marker genes
# res.file<-paste0(workdir, "/analysed_data/differential_analysis_wo2W_BMP4_CNTF_wo2W.xlsx")
# dge_res_sheets <- readxl::excel_sheets(res.file)[-c(1:4)]
# dge_res_sheets <- data.frame(sheets = dge_res_sheets, row.names = dge_res_sheets)
# dge_res_sheets_key <- readxl::read_xlsx(res.file, sheet = "Comparison Key")
# 
# # replace long sheet name from the keys
# dge_res_sheets[dge_res_sheets_key$Key, ] <- dge_res_sheets_key$Comparison
# 
# # select comparisons of interest 
# dgeBulk <- imap(rownames(dge_res_sheets), ~readxl::read_xlsx(res.file, sheet = .x)) %>%
#     setNames(nm = dge_res_sheets$sheets)
# # dgeBulk_sel <- dgeBulk[c("BMP4_diff_3w - progenitor|untreated",
# #                          "CNTF_diff_3w - progenitor|untreated")]
# dgeBulk_sel <- dgeBulk[c(dge_res_sheets_key$Comparison)]
# dgeBulk_sel = dgeBulk_sel[-c(7)]
```

View results table from gene differential analysis.

```{r, message=FALSE, warning=TRUE, fig.height=10, fig.width=10, echo=FALSE, results='asis',fig.keep='all'}
# # apply function to list of comparisons
# clist = 1:length(dgeBulk_sel)
# names(dgeBulk_sel)
# lapply(clist, res_table)
```


```{r}
# countSig <- function(i) { 
#     # select item from list
#     de = dgeBulk_sel[[i]]
#     name=names(dgeBulk_sel)[i]
#     
#     # count significant up and down DEGs
#     deCount = de %>% filter(sig=="<= 0.05") %>% group_by(class) %>% summarise(count=n())
#     
#     return(deCount)
#   }
```

View counts of significant differential analysis hits.

```{r, message=FALSE, warning=TRUE, fig.height=10, fig.width=10, echo=FALSE, results='asis',fig.keep='all'}
# # apply function to list of comparisons
# clist = 1:length(dgeBulk_sel)
# names(dgeBulk_sel)
# lapply(clist, countSig)
```



# Volcano plot: |logFC|>0  {.tabset} 

Visualising the effect sizes and significance for differential gene expression in differentiated and treated astrocytes.
file: differential_analysis_BMP4_CNTF

```{r}
# load de outputs
res.file<-paste0(workdir, "/analysed_data/differential_analysis_BMP4_CNTF.xlsx")
dge_res_sheets <- readxl::excel_sheets(res.file)[-c(1:4)]
dge_res_sheets <- data.frame(sheets = dge_res_sheets, row.names = dge_res_sheets)
dge_res_sheets_key <- readxl::read_xlsx(res.file, sheet = "Comparison Key")

# replace long sheet name from the keys
dge_res_sheets[dge_res_sheets_key$Key, ] <- dge_res_sheets_key$Comparison

# select comparisons of interest 
dgeBulk <- imap(rownames(dge_res_sheets), ~readxl::read_xlsx(res.file, sheet = .x)) %>%
    setNames(nm = dge_res_sheets$sheets)
# dgeBulk_sel <- dgeBulk[c("BMP4_diff_3w - progenitor|untreated",
#                          "CNTF_diff_3w - progenitor|untreated")]
dgeBulk_sel <- dgeBulk[c(dge_res_sheets_key$Comparison)]
# select relevant comparisons
dgeBulk_sel = dgeBulk_sel[-c(6,8,14)]
```


``` {r , echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=3.5, fig.height=2.5, results='asis',fig.keep='all'}

plotl <- c()
for (i in 1:length(dgeBulk_sel)) {
  
    # select relevant comparison
    de = dgeBulk_sel[[i]]
    name=names(dgeBulk_sel)[i]
    # print(name)
    
    # ensure empty gene symbols are replaced with ensembl IDs
    de[is.na(de$symbol),]$symbol <- de[is.na(de$symbol),]$id

    # add a column of NAs
    de$diffexpressed <- NA
    # if log2Foldchange > 0.0 and pvalue < 0.05, set as "UP" 
    de$diffexpressed[de$log2FoldChange > 0.0 & de$pvalue < 0.05] <- "UP"
    # if log2Foldchange < -0.0 and pvalue < 0.05, set as "DOWN"
    de$diffexpressed[de$log2FoldChange < -0.0 & de$pvalue < 0.05] <- "DOWN"
    de$delabel <- NA
    
    # label all up and downregulated genes
    # de[!(is.na(de$diffexpressed)),]$delabel <-  de[!(is.na(de$diffexpressed)),]$symbol
    
    # label top up and downregulated genes
    # largest -log10(p-value)
    lowN = c("BMP4_diff_3w - BMP4_diff_2w|untreated" , "CNTF_diff_3w - CNTF_diff_2w|untreated")
    highN = dge_res_sheets_key$Comparison[-c(3,10)]
    if (name %in%  lowN) {
        N = 5 # number of genes - will depend on the number of significant gene distribution for the DE comparison
    } else {N = 10}
    
    sym = de[head(order(abs(-log10(de$pvalue)), decreasing = TRUE),N),]$symbol 
    de[head(order(abs(-log10(de$pvalue)), decreasing = TRUE),N),]$delabel <- paste0("italic('",sym,"')" )
    # largest log2(FC), also significant
    sigde <- de %>% dplyr::filter(pvalue < 0.05)
    labelid <- sigde[head(order(abs(sigde$log2FoldChange), decreasing = TRUE),7),]$id
    
    sym =  de[de$id %in% labelid,]$symbol
    de[de$id %in% labelid,]$delabel <- paste0("italic('",sym,"')" )
    
    xlim = max(abs(de$log2FoldChange))+(max(abs(de$log2FoldChange))*0.3)
    ylim =  max(-log10(de$pvalue))+(max(-log10(de$pvalue))*0.2)
    
    uplabel = de %>% filter(is.na(delabel)==FALSE) %>% dplyr::filter(diffexpressed=="UP") 
    dwlabel = de %>% filter(is.na(delabel)==FALSE) %>% dplyr::filter(diffexpressed=="DOWN") 

    
    vol_plot <- ggplot(data=de, aes(x=log2FoldChange, y=-log10(pvalue), col=diffexpressed, label=delabel)) +
            geom_point(size = 1, alpha = 0.5) + 
            scale_x_continuous(expand = c(0, 0), limits=c(-xlim,xlim)) +
            scale_y_continuous(expand = c(0, 0), limits=c(0,ylim)) +
            theme_classic(base_size=18) +
            
      geom_label_repel(
                    data = dwlabel,
                    aes(label = delabel),
                     size=6,
                     min.segment.length = unit(0.5, 'lines'), 
                     nudge_y = 5.9,
                     # nudge_x = 7.9,
                     xlim = c(NA, -1),
                     max.overlaps=Inf,
                     parse = TRUE,
                    seed = 1) +
      
      geom_label_repel(
                    data = uplabel,
                    aes(label = delabel),
                     size=6,
                     min.segment.length = unit(0.5, 'lines'),
                     nudge_y = 5.9,
                     # nudge_x = 7.9,
                     xlim = c(1, NA),
                     max.overlaps=Inf,
                     parse = TRUE,
                     seed = 1) +
      
            scale_color_manual(values=c("#788feb", "#942025", "black")) +
            # geom_vline(xintercept=c(-0.6, 0.6), col="black") +
            geom_hline(yintercept=-log10(0.05), col="black", linetype = 2) + 
            ggtitle(name) +
            theme(plot.title = element_text(hjust = 0.5, size=18),
                  legend.position = "none")
    
    # ggsave(file=paste0(workdir,"/volcano_plot/volcano_",name,".svg"), plot=vol_plot, width=6.5, height=6.5)
    
    plotl <- c(plotl, list(vol_plot))
    
    ## extra: for gene-term plot querying
    # de_genes <- de[!is.na(de$delabel),]$delabel
  }

```

```{r, message=FALSE, warning=TRUE, fig.height=10, fig.width=10, echo=FALSE, results='asis',fig.keep='all'}

# names(plotl) <- names(dgeBulk_sel)
# for (i in 1:length(plotl)) {
#   print(plotl[[i]])
# }

for (i in 1:length(dgeBulk_sel)) {

    name = names(dgeBulk_sel)[[i]]
    cat('\n')
    cat('## ', name, '   \n')
    cat('\n')

    print(plotl[[i]])

    cat('\n')

}

```

# Volcano plot: |logFC|>0.5  {.tabset} 

``` {r , echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=3.5, fig.height=2.5, results='asis',fig.keep='all'}

plotl <- c()
for (i in 1:length(dgeBulk_sel)) {
  
    # select relevant comparison
    de = dgeBulk_sel[[i]]
    name=names(dgeBulk_sel)[i]
    # print(name)
    
    # ensure empty gene symbols are replaced with ensembl IDs
    de[is.na(de$symbol),]$symbol <- de[is.na(de$symbol),]$id

    # add a column of NAs
    de$diffexpressed <- NA
    # if log2Foldchange > 0.0 and pvalue < 0.05, set as "UP" 
    de$diffexpressed[de$log2FoldChange > 0.5 & de$pvalue < 0.05] <- "UP"
    # if log2Foldchange < -0.0 and pvalue < 0.05, set as "DOWN"
    de$diffexpressed[de$log2FoldChange < -0.5 & de$pvalue < 0.05] <- "DOWN"
    de$delabel <- NA
    
    # label all up and downregulated genes
    # de[!(is.na(de$diffexpressed)),]$delabel <-  de[!(is.na(de$diffexpressed)),]$symbol
    
    # label top up and downregulated genes
    # largest -log10(p-value)
    lowN = c("BMP4_diff_3w - BMP4_diff_2w|untreated" , "CNTF_diff_3w - CNTF_diff_2w|untreated")
    highN = dge_res_sheets_key$Comparison[-c(3,10)]
    if (name %in%  lowN) {
        N = 5 # number of genes - will depend on the number of significant gene distribution for the DE comparison
    } else {N = 10}
    
    sym = de[head(order(abs(-log10(de$pvalue)), decreasing = TRUE),N),]$symbol 
    de[head(order(abs(-log10(de$pvalue)), decreasing = TRUE),N),]$delabel <- paste0("italic('",sym,"')" )
    # largest log2(FC), also significant
    sigde <- de %>% dplyr::filter(pvalue < 0.05)
    labelid <- sigde[head(order(abs(sigde$log2FoldChange), decreasing = TRUE),7),]$id
    
    sym =  de[de$id %in% labelid,]$symbol
    de[de$id %in% labelid,]$delabel <- paste0("italic('",sym,"')" )
    
    xlim = max(abs(de$log2FoldChange))+(max(abs(de$log2FoldChange))*0.3)
    ylim =  max(-log10(de$pvalue))+(max(-log10(de$pvalue))*0.2)
    
    uplabel = de %>% filter(is.na(delabel)==FALSE) %>% dplyr::filter(diffexpressed=="UP") 
    dwlabel = de %>% filter(is.na(delabel)==FALSE) %>% dplyr::filter(diffexpressed=="DOWN") 

    
    vol_plot <- ggplot(data=de, aes(x=log2FoldChange, y=-log10(pvalue), col=diffexpressed, label=delabel)) +
            geom_point(size = 1, alpha = 0.5) + 
            scale_x_continuous(expand = c(0, 0), limits=c(-xlim,xlim)) +
            scale_y_continuous(expand = c(0, 0), limits=c(0,ylim)) +
            theme_classic(base_size=18) +
            
      geom_label_repel(
                    data = dwlabel,
                    aes(label = delabel),
                     size=6,
                     min.segment.length = unit(0.5, 'lines'), 
                     nudge_y = 5.9,
                     # nudge_x = 7.9,
                     xlim = c(NA, -1),
                     max.overlaps=Inf,
                     parse = TRUE,
                    seed = 1) +
      
      geom_label_repel(
                    data = uplabel,
                    aes(label = delabel),
                     size=6,
                     min.segment.length = unit(0.5, 'lines'),
                     nudge_y = 5.9,
                     # nudge_x = 7.9,
                     xlim = c(1, NA),
                     max.overlaps=Inf,
                     parse = TRUE,
                     seed = 1) +
      
            scale_color_manual(values=c("#788feb", "#942025", "black")) +
            # geom_vline(xintercept=c(-0.6, 0.6), col="black") +
            geom_hline(yintercept=-log10(0.05), col="black", linetype = 2) + 
            ggtitle(name) +
            theme(plot.title = element_text(hjust = 0.5, size=18),
                  legend.position = "none")
    
    # ggsave(file=paste0(workdir,"/volcano_plot/volcano_",name,".svg"), plot=vol_plot, width=6.5, height=6.5)
    
    plotl <- c(plotl, list(vol_plot))
    
    ## extra: for gene-term plot querying
    # de_genes <- de[!is.na(de$delabel),]$delabel
  }

```

```{r, message=FALSE, warning=TRUE, fig.height=10, fig.width=10, echo=FALSE, results='asis',fig.keep='all'}

# names(plotl) <- names(dgeBulk_sel)
# for (i in 1:length(plotl)) {
#   print(plotl[[i]])
# }

for (i in 1:length(dgeBulk_sel)) {

    name = names(dgeBulk_sel)[[i]]
    cat('\n')
    cat('## ', name, '   \n')
    cat('\n')

    print(plotl[[i]])

    cat('\n')

}

```

# Volcano plot: |logFC|>1  {.tabset} 

``` {r , echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=3.5, fig.height=2.5, results='asis',fig.keep='all'}

plotl <- c()
for (i in 1:length(dgeBulk_sel)) {
  
    # select relevant comparison
    de = dgeBulk_sel[[i]]
    name=names(dgeBulk_sel)[i]
    # print(name)
    
    # ensure empty gene symbols are replaced with ensembl IDs
    de[is.na(de$symbol),]$symbol <- de[is.na(de$symbol),]$id

    # add a column of NAs
    de$diffexpressed <- NA
    # if log2Foldchange > 0.0 and pvalue < 0.05, set as "UP" 
    de$diffexpressed[de$log2FoldChange > 1 & de$pvalue < 0.05] <- "UP"
    # if log2Foldchange < -0.0 and pvalue < 0.05, set as "DOWN"
    de$diffexpressed[de$log2FoldChange < -1 & de$pvalue < 0.05] <- "DOWN"
    de$delabel <- NA
    
    # label all up and downregulated genes
    # de[!(is.na(de$diffexpressed)),]$delabel <-  de[!(is.na(de$diffexpressed)),]$symbol
    
    # label top up and downregulated genes
    # largest -log10(p-value)
    lowN = c("BMP4_diff_3w - BMP4_diff_2w|untreated" , "CNTF_diff_3w - CNTF_diff_2w|untreated")
    highN = dge_res_sheets_key$Comparison[-c(3,10)]
    if (name %in%  lowN) {
        N = 5 # number of genes - will depend on the number of significant gene distribution for the DE comparison
    } else {N = 10}
    
    sym = de[head(order(abs(-log10(de$pvalue)), decreasing = TRUE),N),]$symbol 
    de[head(order(abs(-log10(de$pvalue)), decreasing = TRUE),N),]$delabel <- paste0("italic('",sym,"')" )
    # largest log2(FC), also significant
    sigde <- de %>% dplyr::filter(pvalue < 0.05)
    labelid <- sigde[head(order(abs(sigde$log2FoldChange), decreasing = TRUE),7),]$id
    
    sym =  de[de$id %in% labelid,]$symbol
    de[de$id %in% labelid,]$delabel <- paste0("italic('",sym,"')" )
    
    xlim = max(abs(de$log2FoldChange))+(max(abs(de$log2FoldChange))*0.3)
    ylim =  max(-log10(de$pvalue))+(max(-log10(de$pvalue))*0.2)
    
    uplabel = de %>% filter(is.na(delabel)==FALSE) %>% dplyr::filter(diffexpressed=="UP") 
    dwlabel = de %>% filter(is.na(delabel)==FALSE) %>% dplyr::filter(diffexpressed=="DOWN") 

    
    vol_plot <- ggplot(data=de, aes(x=log2FoldChange, y=-log10(pvalue), col=diffexpressed, label=delabel)) +
            geom_point(size = 1, alpha = 0.5) + 
            scale_x_continuous(expand = c(0, 0), limits=c(-xlim,xlim)) +
            scale_y_continuous(expand = c(0, 0), limits=c(0,ylim)) +
            theme_classic(base_size=18) +
            
      geom_label_repel(
                    data = dwlabel,
                    aes(label = delabel),
                     size=6,
                     min.segment.length = unit(0.5, 'lines'), 
                     nudge_y = 5.9,
                     # nudge_x = 7.9,
                     xlim = c(NA, -1),
                     max.overlaps=Inf,
                     parse = TRUE,
                    seed = 1) +
      
      geom_label_repel(
                    data = uplabel,
                    aes(label = delabel),
                     size=6,
                     min.segment.length = unit(0.5, 'lines'),
                     nudge_y = 5.9,
                     # nudge_x = 7.9,
                     xlim = c(1, NA),
                     max.overlaps=Inf,
                     parse = TRUE,
                     seed = 1) +
      
            scale_color_manual(values=c("#788feb", "#942025", "black")) +
            # geom_vline(xintercept=c(-0.6, 0.6), col="black") +
            geom_hline(yintercept=-log10(0.05), col="black", linetype = 2) + 
            ggtitle(name) +
            theme(plot.title = element_text(hjust = 0.5, size=18),
                  legend.position = "none")
    
    ggsave(file=paste0(workdir,"/volcano_plot/volcano_",name,".svg"), plot=vol_plot, width=6.5, height=6.5)
    
    plotl <- c(plotl, list(vol_plot))
    
    ## extra: for gene-term plot querying
    # de_genes <- de[!is.na(de$delabel),]$delabel
  }

```

```{r, message=FALSE, warning=TRUE, fig.height=10, fig.width=10, echo=FALSE, results='asis',fig.keep='all'}

# names(plotl) <- names(dgeBulk_sel)
# for (i in 1:length(plotl)) {
#   print(plotl[[i]])
# }

for (i in 1:length(dgeBulk_sel)) {

    name = names(dgeBulk_sel)[[i]]
    cat('\n')
    cat('## ', name, '   \n')
    cat('\n')

    print(plotl[[i]])

    cat('\n')

}

```

```{r, fig.height=10, fig.width=10}
# apply function to one select comparison
# i=3
# lapply(i, plot_volcano)
```

file: differential_analysis_wo2W_BMP4_CNTF_wo2W

```{r}
# # load differentiated BMP4 and CNTF marker genes
# res.file<-paste0(workdir, "/analysed_data/differential_analysis_wo2W_BMP4_CNTF_wo2W.xlsx")
# dge_res_sheets <- readxl::excel_sheets(res.file)[-c(1:4)]
# dge_res_sheets <- data.frame(sheets = dge_res_sheets, row.names = dge_res_sheets)
# dge_res_sheets_key <- readxl::read_xlsx(res.file, sheet = "Comparison Key")
# 
# # replace long sheet name from the keys
# dge_res_sheets[dge_res_sheets_key$Key, ] <- dge_res_sheets_key$Comparison
# 
# # select comparisons of interest 
# dgeBulk <- imap(rownames(dge_res_sheets), ~readxl::read_xlsx(res.file, sheet = .x)) %>%
#     setNames(nm = dge_res_sheets$sheets)
# # dgeBulk_sel <- dgeBulk[c("BMP4_diff_3w - progenitor|untreated",
# #                          "CNTF_diff_3w - progenitor|untreated")]
# dgeBulk_sel <- dgeBulk[c(dge_res_sheets_key$Comparison)]
# dgeBulk_sel = dgeBulk_sel[-c(7)]
```

``` {r , echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=3.5, fig.height=2.5}

# # create function for plot
# plot_volcano <- function(i) {
#   
#     # select relevant comparison
#     de = dgeBulk_sel[[i]]
#     name=names(dgeBulk_sel)[i]
#     # print(name)
#     
#     # ensure empty gene symbols are replaced with ensembl IDs
#     de[is.na(de$symbol),]$symbol <- de[is.na(de$symbol),]$id
# 
#     # add a column of NAs
#     de$diffexpressed <- NA
#     # if log2Foldchange > 0.0 and pvalue < 0.05, set as "UP" 
#     de$diffexpressed[de$log2FoldChange > 0.0 & de$pvalue < 0.05] <- "UP"
#     # if log2Foldchange < -0.0 and pvalue < 0.05, set as "DOWN"
#     de$diffexpressed[de$log2FoldChange < -0.0 & de$pvalue < 0.05] <- "DOWN"
#     de$delabel <- NA
#     
#     # label all up and downregulated genes
#     # de[!(is.na(de$diffexpressed)),]$delabel <-  de[!(is.na(de$diffexpressed)),]$symbol
#     
#     # label top up and downregulated genes
#     # largest -log10(p-value)
#     lowN = c("BMP4_diff_3w - BMP4_diff_2w|untreated" , "CNTF_diff_3w - CNTF_diff_2w|untreated")
#     highN = dge_res_sheets_key$Comparison[-c(3,10)]
#     if (name %in%  lowN) {
#         N = 5 # number of genes - will depend on the number of significant gene distribution for the DE comparison
#     } else {N = 10}
#     
#     sym = de[head(order(abs(-log10(de$pvalue)), decreasing = TRUE),N),]$symbol 
#     de[head(order(abs(-log10(de$pvalue)), decreasing = TRUE),N),]$delabel <- paste0("italic('",sym,"')" )
#     
#     # largest log2(FC), also significant
#     sigde <- de %>% dplyr::filter(pvalue < 0.05)
#     labelid <- sigde[head(order(abs(sigde$log2FoldChange), decreasing = TRUE),7),]$id
#     sym =  de[de$id %in% labelid,]$symbol
#     de[de$id %in% labelid,]$delabel <- paste0("italic('",sym,"')" )
#     
#     xlim = max(abs(de$log2FoldChange))+(max(abs(de$log2FoldChange))*0.3)
#     ylim = max(-log10(de$pvalue))+(max(-log10(de$pvalue))*0.5)
#     ynud = 
#     
#     uplabel = de %>% filter(is.na(delabel)==FALSE) %>% dplyr::filter(diffexpressed=="UP") 
#     dwlabel = de %>% filter(is.na(delabel)==FALSE) %>% dplyr::filter(diffexpressed=="DOWN") 
# 
#     
#     vol_plot <- ggplot(data=de, aes(x=log2FoldChange, y=-log10(pvalue), col=diffexpressed, label=delabel)) +
#             geom_point(size = 1, alpha = 0.5) + 
#             scale_x_continuous(expand = c(0, 0), limits=c(-xlim,xlim)) +
#             scale_y_continuous(expand = c(0, 0), limits=c(0,ylim)) +
#             theme_classic(base_size=18) +
#             
#       geom_label_repel(
#                     data = dwlabel,
#                     aes(label = delabel),
#                      size=6,
#                      min.segment.length = unit(0.5, 'lines'), 
#                      nudge_y = 5.9,
#                      # nudge_x = 7.9,
#                      xlim = c(NA, -1),
#                      max.overlaps=Inf,
#                      parse = TRUE,
#                     seed = 1) +
#       
#       geom_label_repel(
#                     data = uplabel,
#                     aes(label = delabel),
#                      size=6,
#                      min.segment.length = unit(0.5, 'lines'), 
#                      nudge_y = 5.9,
#                      # nudge_x = 7.9,
#                      xlim = c(1, NA),
#                      max.overlaps=Inf,
#                      parse = TRUE,
#                      seed = 1) +
#       
#             scale_color_manual(values=c("#788feb", "#942025", "black")) +
#             # geom_vline(xintercept=c(-0.6, 0.6), col="black") +
#             geom_hline(yintercept=-log10(0.05), col="black", linetype = 2) + 
#             ggtitle(name) +
#             theme(plot.title = element_text(hjust = 0.5, size=18),
#                   legend.position = "none")
#     
#     # ggsave(file=paste0(workdir,"/volcano_plot/volcano_",name,".svg"), plot=vol_plot, width=5.5, height=5.5)
#     
#     return(vol_plot)
#     
#     ## extra: for gene-term plot querying
#     # de_genes <- de[!is.na(de$delabel),]$delabel
#   }

```

```{r, message=FALSE, warning=TRUE, fig.height=10, fig.width=10, echo=FALSE, results='asis',fig.keep='all'}
# # apply function to list of comparisons
# clist = 1:length(dgeBulk_sel)
# lapply(clist, plot_volcano)
```

```{r, fig.height=10, fig.width=10}
# apply function to one select comparison
# i=3
# lapply(i, plot_volcano)
```

# Venn Diagrams: |logFC|>0  
file: differential_analysis_BMP4_CNTF 
```{r}
# load de outputs
res.file<-paste0(workdir, "/analysed_data/differential_analysis_BMP4_CNTF.xlsx")
dge_res_sheets <- readxl::excel_sheets(res.file)[-c(1:4)]
dge_res_sheets <- data.frame(sheets = dge_res_sheets, row.names = dge_res_sheets)
dge_res_sheets_key <- readxl::read_xlsx(res.file, sheet = "Comparison Key")

# replace long sheet name from the keys
dge_res_sheets[dge_res_sheets_key$Key, ] <- dge_res_sheets_key$Comparison

# select comparisons of interest 
dgeBulk <- imap(rownames(dge_res_sheets), ~readxl::read_xlsx(res.file, sheet = .x)) %>%
    setNames(nm = dge_res_sheets$sheets)
dgeBulk_sel <- dgeBulk[c("BMP4_diff_3w - progenitor|untreated",
                         "CNTF_diff_3w - progenitor|untreated",
                         "treated - untreated|BMP4_diff_3w",
                         "treated - untreated|CNTF_diff_3w")]


select_gene <- function(df) {
  upreg <- df %>% dplyr::filter(pvalue < 0.05) %>% dplyr::filter(log2FoldChange > 0) %>% pull(id)
  downreg <- df %>% dplyr::filter(pvalue < 0.05) %>% dplyr::filter(log2FoldChange < 0) %>% pull(id)
  outdf <- rbind(data.frame(Direction = "Up", id = upreg),
                 data.frame(Direction = "Down", id = downreg))
}
select_gene_outs <- lapply(dgeBulk_sel, select_gene)

##### Differentiated vs Progenitor ##### 
# Upregulated
# List of items
first_up <- select_gene_outs[[1]] %>% dplyr::filter(Direction == "Up") %>% nrow()
second_up <- select_gene_outs[[2]] %>% dplyr::filter(Direction == "Up") %>% nrow()
name1 = paste0("BMP4 v progenitor \n (",first_up,")")
name2 = paste0("CNTF v progenitor \n (",second_up,")")
x <- list( list1 = select_gene_outs[[1]] %>% dplyr::filter(Direction == "Up") %>% pull(id), 
          list2 = select_gene_outs[[2]] %>% dplyr::filter(Direction == "Up") %>% pull(id))
names(x) <- c(name1, name2)
# Plot Venn diagram
p1 <- ggvenn(x, auto_scale = TRUE, text_size = 5, stroke_size = 0.5, set_name_size = 5)
 ggsave(plot=p1, file=paste0(workdir,"/venn_diagram/venn_bmp4_cntf_v_progenitor_upregulated_scaled.svg"), width=12.5, height=12.5)
p2 <- ggvenn(x, auto_scale = FALSE, text_size = 5)
 ggsave(plot=p2, file=paste0(workdir,"/venn_diagram/venn_bmp4_cntf_v_progenitor_upregulated.svg"), width=10, height=12)

# Downregulated
# List of items
first_down <- select_gene_outs[[1]] %>% dplyr::filter(Direction == "Down") %>% nrow()
second_down <- select_gene_outs[[2]] %>% dplyr::filter(Direction == "Down") %>% nrow()
name1 = paste0("BMP4 v progenitor \n (",first_down,")")
name2 = paste0("CNTF v progenitor \n (",second_down,")")
x <- list( list1 = select_gene_outs[[1]] %>% dplyr::filter(Direction == "Down") %>% pull(id), 
          list2 = select_gene_outs[[2]] %>% dplyr::filter(Direction == "Down") %>% pull(id))
names(x) <- c(name1, name2)

# Plot Venn diagram
p3 <- ggvenn(x, auto_scale = TRUE, text_size = 5, stroke_size = 0.5, set_name_size = 5)
 ggsave(plot=p3, file=paste0(workdir,"/venn_diagram/venn_bmp4_cntf_v_progenitor_downregulated_scaled.svg"),  width=12.5, height=12.5)
p4 <- ggvenn(x, auto_scale = FALSE, text_size = 5)
 ggsave(plot=p4, file=paste0(workdir,"/venn_diagram/venn_bmp4_cntf_v_progenitor_downregulated.svg"), width=10, height=12)

##### End of Differentiated vs Progenitor ##### 


 
##### Treated vs Untreated ##### 
# Upregulated
# List of items
first_up <- select_gene_outs[[3]] %>% dplyr::filter(Direction == "Up") %>% nrow()
second_up <- select_gene_outs[[4]] %>% dplyr::filter(Direction == "Up") %>% nrow()
name1 = paste0("treated v untreated BMP4 \n (",first_up,")")
name2 = paste0("treated v untreated CNTF \n (",second_up,")")
x <- list( list1 = select_gene_outs[[3]] %>% dplyr::filter(Direction == "Up") %>% pull(id), 
            list2 = select_gene_outs[[4]] %>% dplyr::filter(Direction == "Up") %>% pull(id))
names(x) <- c(name1, name2)
# Plot Venn diagram
p5 <- ggvenn(x, auto_scale = TRUE, text_size = 5, stroke_size = 0.5, set_name_size = 5)
 ggsave(plot=p5, file=paste0(workdir,"/venn_diagram/venn_treated_v_untreated_upregulated_scaled.svg"), width=12.5, height=12.5)
p6 <- ggvenn(x, auto_scale = FALSE, text_size = 5)
 ggsave(plot=p6, file=paste0(workdir,"/venn_diagram/venn_treated_v_untreated_upregulated.svg"), width=10, height=12)

 
# Downregulated
# List of items
first_down <- select_gene_outs[[3]] %>% dplyr::filter(Direction == "Down") %>% nrow()
second_down <- select_gene_outs[[4]] %>% dplyr::filter(Direction == "Down") %>% nrow()
name1 = paste0("treated v untreated BMP4 \n (",first_down,")")
name2 = paste0("treated v untreated CNTF \n (",second_down,")")
x <- list( list1 = select_gene_outs[[3]] %>% dplyr::filter(Direction == "Down") %>% pull(id), 
            list2 = select_gene_outs[[4]] %>% dplyr::filter(Direction == "Down") %>% pull(id))
names(x) <- c(name1, name2)
# Plot Venn diagram
p7 <- ggvenn(x, auto_scale = TRUE, text_size = 5, stroke_size = 0.5, set_name_size = 5)
 ggsave(plot=p7, file=paste0(workdir,"/venn_diagram/venn_treated_v_untreated_downregulated_scaled.svg"), width=12.5, height=12.5)
p8 <- ggvenn(x, auto_scale = FALSE, text_size = 5)
 ggsave(plot=p8, file=paste0(workdir,"/venn_diagram/venn_treated_v_untreated_downregulated.svg"), width=10, height=12)

##### End of Treated vs Untreated ##### 
 
scplotl <- list(p1,p3,p5,p7) 
nscplotl <- list(p2,p4,p6,p8) 

```

## scaled {.tabset} 
```{r, message=FALSE, warning=TRUE, fig.height=20, fig.width=20, echo=FALSE, results='asis',fig.keep='all'}

plotl = scplotl
names=list("Upregulated Differentiated vs Progenitor", "Downregulated Differentiated vs Progenitor",
        "Upregulated Treated vs Untreated", "Downregulated Treated vs Untreated")

for (i in 1:length(names)) {

    name = names[[i]]
    cat('\n')
    cat('### ', name, '   \n')
    cat('\n')

    print(plotl[[i]])

    cat('\n')

}

```

## non-scaled {.tabset} 
```{r, message=FALSE, warning=TRUE, fig.height=20, fig.width=20, echo=FALSE, results='asis',fig.keep='all'}

plotl = nscplotl
names=list("Upregulated Differentiated vs Progenitor", "Downregulated Differentiated vs Progenitor",
        "Upregulated Treated vs Untreated", "Downregulated Treated vs Untreated")

for (i in 1:length(names)) {

    name = names[[i]]
    cat('\n')
    cat('### ', name, '   \n')
    cat('\n')

    print(plotl[[i]])

    cat('\n')

}

```

# Venn Diagrams: |logFC|>0.5  
file: differential_analysis_BMP4_CNTF 
```{r}
# load de outputs
res.file<-paste0(workdir, "/analysed_data/differential_analysis_BMP4_CNTF.xlsx")
dge_res_sheets <- readxl::excel_sheets(res.file)[-c(1:4)]
dge_res_sheets <- data.frame(sheets = dge_res_sheets, row.names = dge_res_sheets)
dge_res_sheets_key <- readxl::read_xlsx(res.file, sheet = "Comparison Key")

# replace long sheet name from the keys
dge_res_sheets[dge_res_sheets_key$Key, ] <- dge_res_sheets_key$Comparison

# select comparisons of interest 
dgeBulk <- imap(rownames(dge_res_sheets), ~readxl::read_xlsx(res.file, sheet = .x)) %>%
    setNames(nm = dge_res_sheets$sheets)
dgeBulk_sel <- dgeBulk[c("BMP4_diff_3w - progenitor|untreated",
                         "CNTF_diff_3w - progenitor|untreated",
                         "treated - untreated|BMP4_diff_3w",
                         "treated - untreated|CNTF_diff_3w")]


select_gene <- function(df) {
  upreg <- df %>% dplyr::filter(pvalue < 0.05) %>% dplyr::filter(log2FoldChange > 0.5) %>% pull(id)
  downreg <- df %>% dplyr::filter(pvalue < 0.05) %>% dplyr::filter(log2FoldChange < -0.5) %>% pull(id)
  outdf <- rbind(data.frame(Direction = "Up", id = upreg),
                 data.frame(Direction = "Down", id = downreg))
}
select_gene_outs <- lapply(dgeBulk_sel, select_gene)

##### Differentiated vs Progenitor ##### 
# Upregulated
# List of items
first_up <- select_gene_outs[[1]] %>% dplyr::filter(Direction == "Up") %>% nrow()
second_up <- select_gene_outs[[2]] %>% dplyr::filter(Direction == "Up") %>% nrow()
name1 = paste0("BMP4 v progenitor \n (",first_up,")")
name2 = paste0("CNTF v progenitor \n (",second_up,")")
x <- list( list1 = select_gene_outs[[1]] %>% dplyr::filter(Direction == "Up") %>% pull(id), 
          list2 = select_gene_outs[[2]] %>% dplyr::filter(Direction == "Up") %>% pull(id))
names(x) <- c(name1, name2)
# Plot Venn diagram
p1 <- ggvenn(x, auto_scale = TRUE, text_size = 5, stroke_size = 0.5, set_name_size = 5)
 ggsave(plot=p1, file=paste0(workdir,"/venn_diagram/venn_bmp4_cntf_v_progenitor_upregulated_scaled.svg"), width=12.5, height=12.5)
p2 <- ggvenn(x, auto_scale = FALSE, text_size = 5)
 ggsave(plot=p2, file=paste0(workdir,"/venn_diagram/venn_bmp4_cntf_v_progenitor_upregulated.svg"), width=10, height=12)

# Downregulated
# List of items
first_down <- select_gene_outs[[1]] %>% dplyr::filter(Direction == "Down") %>% nrow()
second_down <- select_gene_outs[[2]] %>% dplyr::filter(Direction == "Down") %>% nrow()
name1 = paste0("BMP4 v progenitor \n (",first_down,")")
name2 = paste0("CNTF v progenitor \n (",second_down,")")
x <- list( list1 = select_gene_outs[[1]] %>% dplyr::filter(Direction == "Down") %>% pull(id), 
          list2 = select_gene_outs[[2]] %>% dplyr::filter(Direction == "Down") %>% pull(id))
names(x) <- c(name1, name2)

# Plot Venn diagram
p3 <- ggvenn(x, auto_scale = TRUE, text_size = 5, stroke_size = 0.5, set_name_size = 5)
 ggsave(plot=p3, file=paste0(workdir,"/venn_diagram/venn_bmp4_cntf_v_progenitor_downregulated_scaled.svg"),  width=12.5, height=12.5)
p4 <- ggvenn(x, auto_scale = FALSE, text_size = 5)
 ggsave(plot=p4, file=paste0(workdir,"/venn_diagram/venn_bmp4_cntf_v_progenitor_downregulated.svg"), width=10, height=12)

##### End of Differentiated vs Progenitor ##### 


 
##### Treated vs Untreated ##### 
# Upregulated
# List of items
first_up <- select_gene_outs[[3]] %>% dplyr::filter(Direction == "Up") %>% nrow()
second_up <- select_gene_outs[[4]] %>% dplyr::filter(Direction == "Up") %>% nrow()
name1 = paste0("treated v untreated BMP4 \n (",first_up,")")
name2 = paste0("treated v untreated CNTF \n (",second_up,")")
x <- list( list1 = select_gene_outs[[3]] %>% dplyr::filter(Direction == "Up") %>% pull(id), 
            list2 = select_gene_outs[[4]] %>% dplyr::filter(Direction == "Up") %>% pull(id))
names(x) <- c(name1, name2)
# Plot Venn diagram
p5 <- ggvenn(x, auto_scale = TRUE, text_size = 5, stroke_size = 0.5, set_name_size = 5)
 ggsave(plot=p5, file=paste0(workdir,"/venn_diagram/venn_treated_v_untreated_upregulated_scaled.svg"), width=12.5, height=12.5)
p6 <- ggvenn(x, auto_scale = FALSE, text_size = 5)
 ggsave(plot=p6, file=paste0(workdir,"/venn_diagram/venn_treated_v_untreated_upregulated.svg"), width=10, height=12)

 
# Downregulated
# List of items
first_down <- select_gene_outs[[3]] %>% dplyr::filter(Direction == "Down") %>% nrow()
second_down <- select_gene_outs[[4]] %>% dplyr::filter(Direction == "Down") %>% nrow()
name1 = paste0("treated v untreated BMP4 \n (",first_down,")")
name2 = paste0("treated v untreated CNTF \n (",second_down,")")
x <- list( list1 = select_gene_outs[[3]] %>% dplyr::filter(Direction == "Down") %>% pull(id), 
            list2 = select_gene_outs[[4]] %>% dplyr::filter(Direction == "Down") %>% pull(id))
names(x) <- c(name1, name2)
# Plot Venn diagram
p7 <- ggvenn(x, auto_scale = TRUE, text_size = 5, stroke_size = 0.5, set_name_size = 5)
 ggsave(plot=p7, file=paste0(workdir,"/venn_diagram/venn_treated_v_untreated_downregulated_scaled.svg"), width=12.5, height=12.5)
p8 <- ggvenn(x, auto_scale = FALSE, text_size = 5)
 ggsave(plot=p8, file=paste0(workdir,"/venn_diagram/venn_treated_v_untreated_downregulated.svg"), width=10, height=12)

##### End of Treated vs Untreated ##### 
 
scplotl <- list(p1,p3,p5,p7) 
nscplotl <- list(p2,p4,p6,p8) 

```

## scaled {.tabset} 
```{r, message=FALSE, warning=TRUE, fig.height=20, fig.width=20, echo=FALSE, results='asis',fig.keep='all'}

plotl = scplotl
names=list("Upregulated Differentiated vs Progenitor", "Downregulated Differentiated vs Progenitor",
        "Upregulated Treated vs Untreated", "Downregulated Treated vs Untreated")

for (i in 1:length(names)) {

    name = names[[i]]
    cat('\n')
    cat('### ', name, '   \n')
    cat('\n')

    print(plotl[[i]])

    cat('\n')

}

```

## non-scaled {.tabset} 
```{r, message=FALSE, warning=TRUE, fig.height=20, fig.width=20, echo=FALSE, results='asis',fig.keep='all'}

plotl = nscplotl
names=list("Upregulated Differentiated vs Progenitor", "Downregulated Differentiated vs Progenitor",
        "Upregulated Treated vs Untreated", "Downregulated Treated vs Untreated")

for (i in 1:length(names)) {

    name = names[[i]]
    cat('\n')
    cat('### ', name, '   \n')
    cat('\n')

    print(plotl[[i]])

    cat('\n')

}

```

# Venn Diagrams: |logFC|>1  
file: differential_analysis_BMP4_CNTF 
```{r}
# load de outputs
res.file<-paste0(workdir, "/analysed_data/differential_analysis_BMP4_CNTF.xlsx")
dge_res_sheets <- readxl::excel_sheets(res.file)[-c(1:4)]
dge_res_sheets <- data.frame(sheets = dge_res_sheets, row.names = dge_res_sheets)
dge_res_sheets_key <- readxl::read_xlsx(res.file, sheet = "Comparison Key")

# replace long sheet name from the keys
dge_res_sheets[dge_res_sheets_key$Key, ] <- dge_res_sheets_key$Comparison

# select comparisons of interest 
dgeBulk <- imap(rownames(dge_res_sheets), ~readxl::read_xlsx(res.file, sheet = .x)) %>%
    setNames(nm = dge_res_sheets$sheets)
dgeBulk_sel <- dgeBulk[c("BMP4_diff_3w - progenitor|untreated",
                         "CNTF_diff_3w - progenitor|untreated",
                         "treated - untreated|BMP4_diff_3w",
                         "treated - untreated|CNTF_diff_3w")]


select_gene <- function(df) {
  upreg <- df %>% dplyr::filter(pvalue < 0.05) %>% dplyr::filter(log2FoldChange > 1) %>% pull(id)
  downreg <- df %>% dplyr::filter(pvalue < 0.05) %>% dplyr::filter(log2FoldChange < -1) %>% pull(id)
  outdf <- rbind(data.frame(Direction = "Up", id = upreg),
                 data.frame(Direction = "Down", id = downreg))
}
select_gene_outs <- lapply(dgeBulk_sel, select_gene)

##### Differentiated vs Progenitor ##### 
# Upregulated
# List of items
first_up <- select_gene_outs[[1]] %>% dplyr::filter(Direction == "Up") %>% nrow()
second_up <- select_gene_outs[[2]] %>% dplyr::filter(Direction == "Up") %>% nrow()
name1 = paste0("BMP4 v progenitor \n (",first_up,")")
name2 = paste0("CNTF v progenitor \n (",second_up,")")
x <- list( list1 = select_gene_outs[[1]] %>% dplyr::filter(Direction == "Up") %>% pull(id), 
          list2 = select_gene_outs[[2]] %>% dplyr::filter(Direction == "Up") %>% pull(id))
names(x) <- c(name1, name2)
# Plot Venn diagram
p1 <- ggvenn(x, auto_scale = TRUE, text_size = 5, stroke_size = 0.5, set_name_size = 5)
 ggsave(plot=p1, file=paste0(workdir,"/venn_diagram/venn_bmp4_cntf_v_progenitor_upregulated_scaled.svg"), width=12.5, height=12.5)
p2 <- ggvenn(x, auto_scale = FALSE, text_size = 5)
 ggsave(plot=p2, file=paste0(workdir,"/venn_diagram/venn_bmp4_cntf_v_progenitor_upregulated.svg"), width=10, height=12)

# Downregulated
# List of items
first_down <- select_gene_outs[[1]] %>% dplyr::filter(Direction == "Down") %>% nrow()
second_down <- select_gene_outs[[2]] %>% dplyr::filter(Direction == "Down") %>% nrow()
name1 = paste0("BMP4 v progenitor \n (",first_down,")")
name2 = paste0("CNTF v progenitor \n (",second_down,")")
x <- list( list1 = select_gene_outs[[1]] %>% dplyr::filter(Direction == "Down") %>% pull(id), 
          list2 = select_gene_outs[[2]] %>% dplyr::filter(Direction == "Down") %>% pull(id))
names(x) <- c(name1, name2)

# Plot Venn diagram
p3 <- ggvenn(x, auto_scale = TRUE, text_size = 5, stroke_size = 0.5, set_name_size = 5)
 ggsave(plot=p3, file=paste0(workdir,"/venn_diagram/venn_bmp4_cntf_v_progenitor_downregulated_scaled.svg"),  width=12.5, height=12.5)
p4 <- ggvenn(x, auto_scale = FALSE, text_size = 5)
 ggsave(plot=p4, file=paste0(workdir,"/venn_diagram/venn_bmp4_cntf_v_progenitor_downregulated.svg"), width=10, height=12)

##### End of Differentiated vs Progenitor ##### 


 
##### Treated vs Untreated ##### 
# Upregulated
# List of items
first_up <- select_gene_outs[[3]] %>% dplyr::filter(Direction == "Up") %>% nrow()
second_up <- select_gene_outs[[4]] %>% dplyr::filter(Direction == "Up") %>% nrow()
name1 = paste0("treated v untreated BMP4 \n (",first_up,")")
name2 = paste0("treated v untreated CNTF \n (",second_up,")")
x <- list( list1 = select_gene_outs[[3]] %>% dplyr::filter(Direction == "Up") %>% pull(id), 
            list2 = select_gene_outs[[4]] %>% dplyr::filter(Direction == "Up") %>% pull(id))
names(x) <- c(name1, name2)
# Plot Venn diagram
p5 <- ggvenn(x, auto_scale = TRUE, text_size = 5, stroke_size = 0.5, set_name_size = 5)
 ggsave(plot=p5, file=paste0(workdir,"/venn_diagram/venn_treated_v_untreated_upregulated_scaled.svg"), width=12.5, height=12.5)
p6 <- ggvenn(x, auto_scale = FALSE, text_size = 5)
 ggsave(plot=p6, file=paste0(workdir,"/venn_diagram/venn_treated_v_untreated_upregulated.svg"), width=10, height=12)

 
# Downregulated
# List of items
first_down <- select_gene_outs[[3]] %>% dplyr::filter(Direction == "Down") %>% nrow()
second_down <- select_gene_outs[[4]] %>% dplyr::filter(Direction == "Down") %>% nrow()
name1 = paste0("treated v untreated BMP4 \n (",first_down,")")
name2 = paste0("treated v untreated CNTF \n (",second_down,")")
x <- list( list1 = select_gene_outs[[3]] %>% dplyr::filter(Direction == "Down") %>% pull(id), 
            list2 = select_gene_outs[[4]] %>% dplyr::filter(Direction == "Down") %>% pull(id))
names(x) <- c(name1, name2)
# Plot Venn diagram
p7 <- ggvenn(x, auto_scale = TRUE, text_size = 5, stroke_size = 0.5, set_name_size = 5)
 ggsave(plot=p7, file=paste0(workdir,"/venn_diagram/venn_treated_v_untreated_downregulated_scaled.svg"), width=12.5, height=12.5)
p8 <- ggvenn(x, auto_scale = FALSE, text_size = 5)
 ggsave(plot=p8, file=paste0(workdir,"/venn_diagram/venn_treated_v_untreated_downregulated.svg"), width=10, height=12)

##### End of Treated vs Untreated ##### 
 
scplotl <- list(p1,p3,p5,p7) 
nscplotl <- list(p2,p4,p6,p8) 

```

## scaled {.tabset} 
```{r, message=FALSE, warning=TRUE, fig.height=20, fig.width=20, echo=FALSE, results='asis',fig.keep='all'}

plotl = scplotl
names=list("Upregulated Differentiated vs Progenitor", "Downregulated Differentiated vs Progenitor",
        "Upregulated Treated vs Untreated", "Downregulated Treated vs Untreated")

for (i in 1:length(names)) {

    name = names[[i]]
    cat('\n')
    cat('### ', name, '   \n')
    cat('\n')

    print(plotl[[i]])

    cat('\n')

}

```

## non-scaled {.tabset} 
```{r, message=FALSE, warning=TRUE, fig.height=20, fig.width=20, echo=FALSE, results='asis',fig.keep='all'}

plotl = nscplotl
names=list("Upregulated Differentiated vs Progenitor", "Downregulated Differentiated vs Progenitor",
        "Upregulated Treated vs Untreated", "Downregulated Treated vs Untreated")

for (i in 1:length(names)) {

    name = names[[i]]
    cat('\n')
    cat('### ', name, '   \n')
    cat('\n')

    print(plotl[[i]])

    cat('\n')

}

```

# GO Term Enrichment
file: differential_analysis_BMP4_CNTF
```{r}
# load analysis output
res.file <- list.files(path=paste0(workdir, "/analysed_data/"), pattern="GSEA")
res.df <- data.frame()
for (i in res.file) {
  comparison <- substr(i, 6, nchar(i)-4)
  enr_res <- read.csv(paste0(workdir, "/analysed_data/", i))
  enr_res <- enr_res[,c("gs_subcat","ID","Description","enrichmentScore","NES","p.adjust","core_enrichment","core_enrichment_shrunkLFC")]
  res.df <- rbind(res.df, data.frame(Comparison = comparison, 
                                     enr_res))
}

# review loaded data
length(table(res.df$Comparison)) # 7 comparisons
data.frame(table(res.df$Comparison)) # Number of GO terms by comparison
table(res.df$gs_subcat) # 1 database

# # edit descriptions
# res.df$Description <- substr(res.df$Description,6, nchar(res.df$Description))
res.df$Description <- gsub("_", " ", res.df$Description)

## make description word lowercase
# first_l <- substr(res.df$Description,1, 1)
# second_l <- tolower(substr(res.df$Description, 2, nchar(res.df$Description)))
# all_l <- paste0(first_l, second_l)
# res.df$Description <- all_l

# annotate positive and negative NES
res.df$nes_sym <- ifelse( res.df$NES < 0, "neg", "pos")

```


```{r, fig.height=16, fig.width=16}
# plot figures
dirs = unique(res.df$nes_sym)
comps = unique(res.df$Comparison) %>% 
  grep("treated - untreated_in_CNTF_diff_3w - BMP4_diff_3w", ., invert=TRUE, value=TRUE) # filter out unused comparison

# plotting function
plotORA <- function(results) {
  gnps <- brewer.pal(9, 'GnBu')
  gnps <- gnps[c(5:9)]
  results$Description <- as.character(results$Description) %>% str_wrap(., width=60)
  gsea_plot <- ggplot(results, aes(x = NES, y = reorder(Description, abs(NES)), fill = -log10(p.adjust))) +
  geom_bar(stat = "identity") +
  scale_fill_gradientn(colours=gnps) +
  labs(title = paste0(comparison," ",direction_name),
       x = "NES",
       y = "GO:BP Gene set",
       color = "-log10(p-value)") +
  theme_bw(base_size = 18) +
  theme(plot.title = element_text(hjust = 0.5) # Center the plot title
        ) + 
  scale_y_discrete(labels=function(x) str_wrap(x, width=60)) 
  return(gsea_plot)
}

# loop function over all groups
plotList <- c()
nameList <- c()
for (comparison in comps) {
  for (direction in dirs) {
    direction_name = ifelse(direction=="neg","downregulated",ifelse(direction=="pos","upregulated","NA"))
    # filter the top 50 GO terms
    res_filter <- res.df %>%
                  dplyr::filter(Comparison == comparison) %>%
                  dplyr::filter(nes_sym == direction) %>%
                  arrange(desc(abs(NES))) %>%
                  slice_head(n = 50)
    
    plot <- plotORA(res_filter)
    name <- paste0(comparison, "_", direction_name)
    plotList <- c(plotList, list(plot))
    nameList <- c(nameList, list(name))
    # ggsave(file=paste0(workdir,"/gsea_plot/gsea_",comparison,"_",direction_name,"_top50_gobp.svg"), plot=gsea_plot, width=18, height=25)
  }
}
names(plotList) <- nameList
```

```{r}
# save table
dirs = unique(res.df$nes_sym)
comps = unique(res.df$Comparison)

tableList <- c()
nameList <- c()
for (comparison in comps) {
  for (direction in dirs) {
    direction_name = ifelse(direction=="neg","downregulated",ifelse(direction=="pos","upregulated","NA"))
    # filter the top 50 GO terms
    res_filter <- res.df %>%
                  dplyr::filter(Comparison == comparison) %>%
                  dplyr::filter(nes_sym == direction) %>%
                  arrange(desc(abs(NES))) %>%
                  slice_head(n = 50)
    res_slim = res_filter[c("ID","enrichmentScore","NES","p.adjust","core_enrichment","core_enrichment_shrunkLFC")]
    # save extended file
    savedir="/Users/k23139900/Documents/PhD/projects/astrocyte_differentiation/pu_analysis/GOtermtables"
    write.csv(res_filter,paste0(savedir,"/",comparison,"_",direction_name,"_top50_gobp_extended.csv"))
    
    # save slim file
    res_slim = res_filter[c("ID","NES","p.adjust")]
    savedir="/Users/k23139900/Documents/PhD/projects/astrocyte_differentiation/pu_analysis/GOtermtables"
    write.csv(res_slim,paste0(savedir,"/",comparison,"_",direction_name,"_top50_gobp.csv"))
    
    name <- paste0(comparison,"_",direction_name)
    tableList <- c(tableList, list(res_slim))
    nameList <- c(nameList, list(name))
  }
}

names(tableList) <- nameList
```

## plots {.tabset} 
```{r, message=FALSE, warning=TRUE, fig.height=20, fig.width=18, echo=FALSE, results='asis',fig.keep='all'}

plotl = plotList
names = names(plotList)

for (i in 1:length(names)) {

    name = names[[i]]
    cat('\n')
    cat('### ', name, '   \n')
    cat('\n')

    print(plotl[[i]])

    cat('\n')

}

```

# Session Info
```{r}
sessionInfo()
```