---
title: "siletti_marker_enrichment_figure"
author: "Paulina"
date: "2024-08-30"
output:
  rmarkdown::html_document:
    theme: default
    highlight: tango
    code_folding: show
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
    number_sections: no
    self_contained: yes
    df_print: paged
---

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DESeq2)
library(ggpubr)
library(BPCells) # remotes::install_github("bnprks/BPCells")
library(Seurat) 
library(SeuratObject)
library(SeuratDisk)
library(Azimuth) # remotes::install_github('satijalab/azimuth', ref = 'master')
library(circlize)
library(purrr)
library(stringr)
library(tidyr)
library(dplyr)
library(ComplexHeatmap)
```

``` {r}
workdir = "/home/pu1918/RDS/user/pu1918/ephemeral/AS_astrocytes"
```


Reference dataset (Siletti et al. 2023) can be downloaded from CELLXGENE (Human Brain Cell Atlas v1.0)
``` {bash}
# 3 Files: All neurons h5ad  (31G); All non-neurons h5ad (4.4G); Supercluster: Astrocyte (850M)
# wget https://datasets.cellxgene.cziscience.com/b9171f05-8112-4a55-95f2-4cf8a57df8a2.h5ad
# wget https://datasets.cellxgene.cziscience.com/cc9bfb86-96ed-4ecd-bcc9-464120fc8628.h5ad
# wget https://datasets.cellxgene.cziscience.com/f86dbeff-36bf-4036-a828-c2c43b052cf2.rds
```

```{r Function to format count assay}
## function to be used when data assay contains raw counts. This replaces the
## information into counts assay, and normalises counts for the data assay.
## There's also an option to replace gene symbols as matrix rownames.
formatassay <- function(obj) {
  # extract data from seurat object
  assay <- obj[['RNA']]@data

  # gene information (rows)
  metafeature <- obj[['RNA']]@meta.features
  metafeature$ensembl_id <- row.names(metafeature)
  head(metafeature)

  # Option to switch ENSEMBL IDs to gene symbols
  #   # where gene names repeat in a column, and you want to add a unique number to the end of each gene name
  #   metafeature$Gene <- as.character(metafeature$Gene)
  #   metafeature$Gene <- ave(metafeature$Gene, metafeature$Gene, FUN = function(x) {
  #     if(length(x) > 1) {
  #       letters_combinations <- c(letters, sapply(letters, function(l1) paste0(letters, l1)))
  #       suffix <- letters_combinations[seq_along(x)]
  #       paste0(x, "_", suffix)
  #     } else {
  #       x
  #     }
  #   })
  #
  # rownames(metafeature) <- metafeature$Gene
  # rownames(assay) <- metafeature$Gene

  # cell information (columns)
  metadata <- obj@meta.data
  # dimensionality reductions
  red_umap <- obj$UMAP
  red_tsne <- obj$tSNE
  # create new object
  newobj <- CreateSeuratObject(counts = assay, meta.data = metadata)
  # stilobj[["RNA"]]@meta.features <- metafeature # ‘meta.features’ is not a slot in class “Assay5”
  newobj$UMAP <- red_umap
  newobj$tSNE <- red_tsne
  # normalize counts by cell (scale factor from total reads per cell)
  newobj <- NormalizeData(newobj)
  # rename
  obj <- newobj
  # check
  obj
  head(rownames(obj))
  # drop large unused objects
  newobj <- NULL
  assay <- NULL

  return(obj)
}
```

```{r Create reference RDS object}
# ### Create reference object with ENSEMBL IDs as rownames
# 
# # open h5ad files using BPCells (for large files)
# file.dir <- paste0(workdir, "/ref_siletti/")
# files.set <- c("b9171f05-8112-4a55-95f2-4cf8a57df8a2.h5ad","cc9bfb86-96ed-4ecd-bcc9-464120fc8628.h5ad") # Non-neurons, Neurons
# 
# # Loop through (2) h5ad files and output BPCells matrices on-disk
# data.list <- c()
# metadata.list <- c()
# for (i in 1:length(files.set)) {
#   path <- paste0(file.dir, files.set[i])
#   data <- BPCells::open_matrix_anndata_hdf5(path, group = "X")
#    write_matrix_dir(
#      mat = data,
#      dir = paste0(gsub(".h5ad", "", path), "_BP")
#    )
#   # Load in BP matrices
#   mat <- open_matrix_dir(dir = paste0(gsub(".h5ad", "", path), "_BP"))
#   # Get metadata
#   metadata.list[[i]] <- LoadH5ADobs(path = path)
#   data.list[[i]] <- mat
# }
# # Name layers
# names(data.list) <- c("nonneuron", "neuron")
# 
# # check metadata columns
# colnames(metadata.list[[i]])
# # Add Metadata
# metadata.list <- lapply(metadata.list, function(x) {
#   x <- x[, c("CellID", "cell_cycle_score", "cell_type", "cluster_id", "development_stage",
#              "disease","donor_id","fraction_mitochondrial", "fraction_unspliced" ,
#             "sample_id","self_reported_ethnicity", "sex", "supercluster_term",
#             "total_UMIs", "total_genes")]
#   return(x)
# })
# metadata <- Reduce(rbind, metadata.list)
# 
# # create seurat object for both datasets
# merged.object <- CreateSeuratObject(counts = data.list, meta.data = metadata) # 59236 features across 3369219 samples within 1 assay
# # save object
# saveRDS(
#   object = merged.object,
#   file = paste0(workdir, "/ref_siletti/ref_siletti_neuron_nonneuron_seurat_supercluster_term_ensembl.Rds")
# )
# # drop
# merged.object<-NULL
```

# Human astrocyte cell type marker gene enrichment
## Load data
```{r load seurat object}
# load RDS data file
file="ref_siletti_neuron_nonneuron_seurat_supercluster_term_ensembl.Rds"       # ref_siletti_neuron_nonneuron_seurat_supercluster_term_ensembl.Rds OR 
                                                      # "f86dbeff-36bf-4036-a828-c2c43b052cf2.rds" (all neurons + non-neurons, or astrocytes only)
seu <- readRDS(paste0(workdir, "/ref_siletti/",file))
seu
```


``` {r subtype cell count by cluster id, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=7.5, fig.height=5.25}
seu <- SetIdent(seu, value = "cell_type")
# count cells per cluster
cellcount <- data.frame(table(Idents(seu)))
ggplot(cellcount, aes(x=Var1, y=Freq)) + geom_bar( stat="identity") + 
  theme_bw() + ylab("Cell counts") + xlab("Cluster") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
# ggsave(file=paste0(workdir,"/ref_stiletti/plots/astrocyte_celltype_cellcounts_cluster_id.svg"), width=7.5, height=3.5)
```

``` {r subtype cell count by donor id, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=4.5, fig.height=2.25}
seu <- SetIdent(seu, value = "donor_id")
# count cells per cluster
cellcount <- data.frame(table(Idents(seu)))
ggplot(cellcount, aes(x=Var1, y=Freq)) + geom_bar( stat="identity") + 
  theme_bw() + ylab("Cell counts") + xlab("Donor")
# ggsave(file=paste0(workdir,"/ref_stiletti/plots/astrocyte_celltype_cellcounts_donor_id.svg"), width=7.5, height=3.5)

```

## Find marker genes

```{r find subtype marker}
# find upregulated markers for all astrocyte subtype clusters versus all other cell types
seu <- SetIdent(seu, value = "cell_type")
seu.markers <- FindAllMarkers(seu,
                                  group.by = "cell_type",
                                  only.pos = TRUE,
                                  min.pct = 0.1,
                                  logfc.threshold = 0.25,
                                  test.use = "wilcox")
write.table(seu.markers, paste0(workdir,"/ref_siletti/stiletti_astrocyte_celltype_singlecell_wilcox_minpct0.1_marker_genes.txt"))
```

```{r}
# read in marker gene list
stilSeu.markers <- read.table(paste0(workdir,"/ref_siletti/stiletti_astrocyte_celltype_singlecell_wilcox_minpct0.1_marker_genes.txt")) 
```

View top 10 markers for each astrocyte subtype
```{r}
# view top 10 markers by subtype
stilSeu.markers %>% group_by(cluster) %>% slice_head(n=10)
```


Extract top 50 markers for each astrocyte subtype
```{r}
# filter by significance and sort by effect size, and subset to top N markers
sN = 20
stilSeu.markers <- stilSeu.markers %>%
                  dplyr::filter(p_val_adj < 0.05) %>%
                  group_by(cluster) %>%
                  arrange(desc(abs(avg_log2FC))) %>% # abs() not needed since have selected only upregulated genes
                  # arrange(p_val_adj) %>% # does not really work when so many p-values are 0s
                  slice_head(n = sN) %>%
                  ungroup()

# check
stilSeu.markers %>% group_by(cluster) %>% summarize(N=n())
```

Find marker genes among top bulk DEGs
```{r}
# load differentiated BMP4 and CNTF marker genes
res.file <- paste0(workdir, "/analysed_data/differential_analysis_wo2W_BMP4_CNTF_wo2W.xlsx")
dge_res_sheets <- readxl::excel_sheets(res.file)[-c(1:4)]
dge_res_sheets <- data.frame(sheets = dge_res_sheets, row.names = dge_res_sheets)
dge_res_sheets_key <- readxl::read_xlsx(res.file, sheet = "Comparison Key")
# Replace long sheet name from the keys
dge_res_sheets[dge_res_sheets_key$Key, ] <- dge_res_sheets_key$Comparison
# select comparisons of interest 
dgeBulk <- imap(rownames(dge_res_sheets), ~readxl::read_xlsx(res.file, sheet = .x)) %>%
    setNames(nm = dge_res_sheets$sheets)
dgeBulk_sel <- dgeBulk[c("BMP4_diff_3w - progenitor|untreated",
                         "CNTF_diff_3w - progenitor|untreated")]

```

```{r}
# find common genes 
# bN = 200
bN = 500                                   # edit to test different number of differential genes from bulk RNA dataset
outs_l <- list()
for (i in 1:length(dgeBulk_sel)) {
  
  bulkRNA = dgeBulk_sel[[i]]
  bulkMark <- bulkRNA %>% 
    filter(!(sig=="NS")) %>% 
    filter(!(class=="Down*")) %>%           # only upregulated marker genes
    arrange(desc(abs(log2FoldChange))) %>% 
    slice_head(n = bN)                      
  outdf <- merge(bulkMark, stilSeu.markers, by.x="id", by.y="gene")
  outdf$Comparison <- names(dgeBulk_sel)[i]
  outs_l <- c(outs_l, list(outdf))
}
names(outs_l)<-names(dgeBulk_sel)
outs_df <- do.call("rbind",outs_l)
outs_df

outs_df %>% group_by(cluster) %>% slice_head(n=10)
```

Number of overlapping marker genes by astrocyte differentiation
```{r}
# add number of used cluster marker genes 
if (is.null( outs_df$top.prop)  ) 
{outs_df$top.prop <- sN}

# number of overlapping marker genes by astrocyte differentiation
ol_count <- data.frame(table(outs_df$Comparison))
colnames(ol_count) <- c("Comparison","OLCount")
# ol_count$scProp <- ol_count$OLCount / sN
# ol_count$bProp <- ol_count$OLCount / bN
ol_count
```

Number of overlapping genes by cluster
```{r}
# get number of overlapping genes by cluster
cluster_ol_gene_count <- outs_df %>% 
  group_by(Comparison, cluster) %>% 
  summarise(N.overlap=n()) %>% 
  as.data.frame()
# add number of cluster marker genes tested for overlap (top.prop)
cluster_ol_gene_count <- merge(cluster_ol_gene_count, unique(outs_df[,c("Comparison","cluster","top.prop")]), by = c("Comparison","cluster"))
cluster_ol_gene_count$GeneRatio <- cluster_ol_gene_count$N.overlap / cluster_ol_gene_count$top.prop
```

```{r}
# apply hypergeometric test for overlap
backpop = 18372               # total N genes
outdf = data.frame()
for (i in 1:nrow(cluster_ol_gene_count)) {
  test_row = cluster_ol_gene_count[i,]
  clustermark = test_row$top.prop
  overlap = test_row$N.overlap
  
  signature_l = c(bN,bN)    # N genes limited by bN threshold
  # signature_l = c(3290,3278)  # N significant bulk DEGs
  signature = ifelse(test_row$Comparison == "BMP4_diff_3w - progenitor|untreated", signature_l[1], signature_l[2])
  
  pval = phyper(q=overlap-1,            # number of overlapping genes - 1
                m=clustermark,          # number of cluster marker genes
                n=backpop-clustermark,  # number of genes tested - overlapping genes
                k=signature,            # number of differentiated astrocyte DEGs
                lower.tail=FALSE)       # compute P( X > overlap ), hence the '-1' above
  out <- data.frame(test_row, pval=pval)
  outdf <- rbind(outdf, out)
}

# apply multiple testing correction
outdf$p.adjust <- p.adjust(outdf$pval, method = "fdr")
```

```{r}
outdf
```

```{r, fig.height=5, fig.width=10}
# rename clusters
outdf$cluster <- paste0("Astro_",outdf$cluster)

# rename comparisons
outdf[outdf$Comparison == c("BMP4_diff_3w - progenitor|untreated"),]$Comparison <- "3w BMP4\nastrocyte"
outdf[outdf$Comparison == c("CNTF_diff_3w - progenitor|untreated"),]$Comparison <- "3w CNTF\nastrocyte"

# plot counts
outdf$cluster <- as.factor(outdf$cluster)
ggplot(outdf, aes(x=cluster, y=N.overlap, fill=Comparison, label = ifelse(p.adjust < 0.05, 
               paste0("p = ",formatC(p.adjust, format = "e", digits = 2)), 
               ""))) +
  geom_text(position=position_dodge(width = 0.9), vjust = -0.5, size=3) +
  geom_bar(position = "dodge", stat="identity") + 
  theme_bw() +
  ggtitle("Common Human Brain Atlas astrocyte celltype and differentiated astrocyte marker genes")
# ggsave(file=paste0(workdir,"/ref_stiletti/plots/barplot_bmp4_cntf_v_progenitor_upregulated.svg"), width=16, height=5)

# plot gene ratios
outdf$cluster <- as.factor(outdf$cluster)
ggplot(outdf, aes(x=cluster, y=GeneRatio, fill=Comparison, label = ifelse(p.adjust < 0.05, 
               paste0("p = ",formatC(p.adjust, format = "e", digits = 2)), 
               ""))) +
  geom_text(position=position_dodge(width = 0.9), vjust = -0.5, size=3) +
  geom_bar(position = "dodge", stat="identity") + 
  theme_bw() +
  ggtitle("Common Human Brain Atlas astrocyte celltype and differentiated astrocyte marker genes")
```

```{r, fig.height=4, fig.width=8}
# Plot heamap of ratios
ggplot(outdf, aes(cluster, Comparison, fill= GeneRatio)) + 
  geom_tile() +
  scale_fill_gradient(low="white", high="darkgreen") +
  # geom_text(aes(label = ifelse(p.adjust < 0.05, 
  #                              paste0("p = ",formatC(p.adjust, format = "e", digits = 2)), 
  #                              "")), size=2.5)
  geom_text(aes(label = ifelse( p.adjust < 0.0005, "***",
                                ifelse(p.adjust < 0.05, "**", 
                                       ifelse(p.adjust < 0.1, "*","")))), size=5) +
  theme_bw() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) 
  
# ggsave(file=paste0(dir,"/ref_stiletti/plots/heatmap_bmp4_cntf_siletti_celltype_marker_enrich.svg"), width=16, height=5)
# ggsave(file=paste0(dir,"/ref_stiletti/plots/heatmap_bmp4_cntf_siletti_celltype_marker_enrich.png"), width=9, height=3)

```



# Human astrocyte subtype marker gene enrichment
## Load data
```{r load seurat object}
# load RDS data file
file="f86dbeff-36bf-4036-a828-c2c43b052cf2.rds"       # ref_siletti_neuron_nonneuron_seurat_supercluster_term_ensembl.Rds OR 
                                                      # "f86dbeff-36bf-4036-a828-c2c43b052cf2.rds" (all neurons + non-neurons, or astrocytes only)
seu <- readRDS(paste0(workdir, "/ref_siletti/",file))
seu
```

```{r}
# format data
seu <- formatassay(seu)
seu
```

``` {r subtype cell count by cluster id, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=4.5, fig.height=2.25}
seu <- SetIdent(seu, value = "cluster_id")
# count cells per cluster
cellcount <- data.frame(table(Idents(seu)))
ggplot(cellcount, aes(x=Var1, y=Freq)) + geom_bar( stat="identity") + 
  theme_bw() + ylab("Cell counts") + xlab("Cluster")
# ggsave(file=paste0(workdir,"/ref_stiletti/plots/astrocyte_supercluster_cellcounts_cluster_id.svg"), width=7.5, height=3.5)
```

``` {r subtype cell count by donor id, echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=4.5, fig.height=2.25}
seu <- SetIdent(seu, value = "donor_id")
# count cells per cluster
cellcount <- data.frame(table(Idents(seu)))
ggplot(cellcount, aes(x=Var1, y=Freq)) + geom_bar( stat="identity") + 
  theme_bw() + ylab("Cell counts") + xlab("Donor")
# ggsave(file=paste0(workdir,"/ref_stiletti/plots/astrocyte_supercluster_cellcounts_donor_id.svg"), width=7.5, height=3.5)

```

## Find marker genes

```{r find subtype marker}
# # find upregulated markers for all astrocyte subtype clusters versus all other cell types
# seu <- SetIdent(seu, value = "cluster_id")
# seu.markers <- FindAllMarkers(seu, 
#                                   group.by = "cluster_id",
#                                   only.pos = TRUE, 
#                                   min.pct = 0.1, 
#                                   logfc.threshold = 0.25,
#                                   test.use = "wilcox")
# write.table(seu.markers, paste0(workdir,"/ref_siletti/stiletti_astrocyte_subcluster_singlecell_wilcox_minpct0.1_marker_genes.txt"))
```

```{r}
# read in marker gene list
stilSeu.markers <- read.table(paste0(workdir,"/ref_siletti/stiletti_astrocyte_subcluster_singlecell_wilcox_minpct0.1_marker_genes.txt")) 
```

View top 10 markers for each astrocyte subtype
```{r}
# view top 10 markers by subtype
stilSeu.markers %>% group_by(cluster) %>% slice_head(n=10)
```


Extract top 50 markers for each astrocyte subtype
```{r}
# filter by significance and sort by effect size, and subset to top N markers
sN = 20
stilSeu.markers <- stilSeu.markers %>%
                  dplyr::filter(p_val_adj < 0.05) %>%
                  group_by(cluster) %>%
                  arrange(desc(abs(avg_log2FC))) %>% # abs() not needed since have selected only upregulated genes
                  # arrange(p_val_adj) %>% # does not really work when so many p-values are 0s
                  slice_head(n = sN) %>%
                  ungroup()

# check
stilSeu.markers %>% group_by(cluster) %>% summarize(N=n())
```

Find marker genes among top bulk DEGs
```{r}
# load differentiated BMP4 and CNTF marker genes
res.file <- paste0(workdir, "/analysed_data/differential_analysis_wo2W_BMP4_CNTF_wo2W.xlsx")
dge_res_sheets <- readxl::excel_sheets(res.file)[-c(1:4)]
dge_res_sheets <- data.frame(sheets = dge_res_sheets, row.names = dge_res_sheets)
dge_res_sheets_key <- readxl::read_xlsx(res.file, sheet = "Comparison Key")
# Replace long sheet name from the keys
dge_res_sheets[dge_res_sheets_key$Key, ] <- dge_res_sheets_key$Comparison
# select comparisons of interest 
dgeBulk <- imap(rownames(dge_res_sheets), ~readxl::read_xlsx(res.file, sheet = .x)) %>%
    setNames(nm = dge_res_sheets$sheets)
dgeBulk_sel <- dgeBulk[c("BMP4_diff_3w - progenitor|untreated",
                         "CNTF_diff_3w - progenitor|untreated")]

```

```{r}
# find common genes 
# bN = 200
bN = 500                                   # edit to test different number of differential genes from bulk RNA dataset
outs_l <- list()
for (i in 1:length(dgeBulk_sel)) {
  
  bulkRNA = dgeBulk_sel[[i]]
  bulkMark <- bulkRNA %>% 
    filter(!(sig=="NS")) %>% 
    filter(!(class=="Down*")) %>%           # only upregulated marker genes
    arrange(desc(abs(log2FoldChange))) %>% 
    slice_head(n = bN)                      
  outdf <- merge(bulkMark, stilSeu.markers, by.x="id", by.y="gene")
  outdf$Comparison <- names(dgeBulk_sel)[i]
  outs_l <- c(outs_l, list(outdf))
}
names(outs_l)<-names(dgeBulk_sel)
outs_df <- do.call("rbind",outs_l)
outs_df

outs_df %>% group_by(cluster) %>% slice_head(n=10)
```

Number of overlapping marker genes by astrocyte differentiation
```{r}
# add number of used cluster marker genes 
if (is.null( outs_df$top.prop)  ) 
{outs_df$top.prop <- sN}

# number of overlapping marker genes by astrocyte differentiation
ol_count <- data.frame(table(outs_df$Comparison))
colnames(ol_count) <- c("Comparison","OLCount")
# ol_count$scProp <- ol_count$OLCount / sN
# ol_count$bProp <- ol_count$OLCount / bN
ol_count
```

Number of overlapping genes by cluster
```{r}
# get number of overlapping genes by cluster
cluster_ol_gene_count <- outs_df %>% 
  group_by(Comparison, cluster) %>% 
  summarise(N.overlap=n()) %>% 
  as.data.frame()
# add number of cluster marker genes tested for overlap (top.prop)
cluster_ol_gene_count <- merge(cluster_ol_gene_count, unique(outs_df[,c("Comparison","cluster","top.prop")]), by = c("Comparison","cluster"))
cluster_ol_gene_count$GeneRatio <- cluster_ol_gene_count$N.overlap / cluster_ol_gene_count$top.prop
```

```{r}
# apply hypergeometric test for overlap
backpop = 18372               # total N genes
outdf = data.frame()
for (i in 1:nrow(cluster_ol_gene_count)) {
  test_row = cluster_ol_gene_count[i,]
  clustermark = test_row$top.prop
  overlap = test_row$N.overlap
  
  signature_l = c(bN,bN)    # N genes limited by bN threshold
  # signature_l = c(3290,3278)  # N significant bulk DEGs
  signature = ifelse(test_row$Comparison == "BMP4_diff_3w - progenitor|untreated", signature_l[1], signature_l[2])
  
  pval = phyper(q=overlap-1,            # number of overlapping genes - 1
                m=clustermark,          # number of cluster marker genes
                n=backpop-clustermark,  # number of genes tested - overlapping genes
                k=signature,            # number of differentiated astrocyte DEGs
                lower.tail=FALSE)       # compute P( X > overlap ), hence the '-1' above
  out <- data.frame(test_row, pval=pval)
  outdf <- rbind(outdf, out)
}

# apply multiple testing correction
outdf$p.adjust <- p.adjust(outdf$pval, method = "fdr")
```

```{r}
outdf
```

```{r, fig.height=5, fig.width=10}
# rename clusters
outdf$cluster <- paste0("Astro_",outdf$cluster)

# rename comparisons
outdf[outdf$Comparison == c("BMP4_diff_3w - progenitor|untreated"),]$Comparison <- "3w BMP4\nastrocyte"
outdf[outdf$Comparison == c("CNTF_diff_3w - progenitor|untreated"),]$Comparison <- "3w CNTF\nastrocyte"

# plot counts
outdf$cluster <- as.factor(outdf$cluster)
ggplot(outdf, aes(x=cluster, y=N.overlap, fill=Comparison, label = ifelse(p.adjust < 0.05, 
               paste0("p = ",formatC(p.adjust, format = "e", digits = 2)), 
               ""))) +
  geom_text(position=position_dodge(width = 0.9), vjust = -0.5, size=3) +
  geom_bar(position = "dodge", stat="identity") + 
  theme_bw() +
  ggtitle("Common Human Brain Atlas astrocyte subcluster and differentiated astrocyte marker genes")
# ggsave(file=paste0(workdir,"/ref_stiletti/plots/barplot_bmp4_cntf_v_progenitor_upregulated.svg"), width=16, height=5)

# plot gene ratios
outdf$cluster <- as.factor(outdf$cluster)
ggplot(outdf, aes(x=cluster, y=GeneRatio, fill=Comparison, label = ifelse(p.adjust < 0.05, 
               paste0("p = ",formatC(p.adjust, format = "e", digits = 2)), 
               ""))) +
  geom_text(position=position_dodge(width = 0.9), vjust = -0.5, size=3) +
  geom_bar(position = "dodge", stat="identity") + 
  theme_bw() +
  ggtitle("Common Human Brain Atlas astrocyte subcluster and differentiated astrocyte marker genes")
```

```{r, fig.height=4, fig.width=8}
# Plot heamap of ratios
ggplot(outdf, aes(cluster, Comparison, fill= GeneRatio)) + 
  geom_tile() +
  scale_fill_gradient(low="white", high="darkgreen") +
  # geom_text(aes(label = ifelse(p.adjust < 0.05, 
  #                              paste0("p = ",formatC(p.adjust, format = "e", digits = 2)), 
  #                              "")), size=2.5)
  geom_text(aes(label = ifelse( p.adjust < 0.0005, "***",
                                ifelse(p.adjust < 0.05, "**", 
                                       ifelse(p.adjust < 0.1, "*","")))), size=5) +
  theme_bw() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) 
  
# ggsave(file=paste0(dir,"/ref_stiletti/plots/heatmap_bmp4_cntf_siletti_subcluster_marker_enrich.svg"), width=16, height=5)
# ggsave(file=paste0(dir,"/ref_stiletti/plots/heatmap_bmp4_cntf_siletti_subcluster_marker_enrich.png"), width=9, height=3)

```



```{r}
sessionInfo()
```